\documentclass[12pt]{letter}

\usepackage[utf8]{inputenc}
\usepackage[margin=1in]{geometry}


\begin{document}

\date{November 20, 2024}

\begin{letter}{Professor João Soares \\
Editor-in-Chief \\
Canadian Journal of Chemical Engineering}

\opening{Dear Professor João Soares,}

Thank you for considering our manuscript, \textit{Optimal Control of Axial Dispersion Tubular Reactors with Recycle: Addressing State-Delay through Transport PDEs}, for publication. We are also grateful to the reviewers for their valuable comments and suggestions, which have been very helpful in improving our work.

To assist in the review process, we have included the following:
\begin{itemize}
    \item \textbf{Reply letter}: Following this cover letter, detailed responses to each reviewer’s comment is provided. In addition to addressing each comment individually, we have also discussed two common concerns raised by multiple reviewers at the beginning of the reply letter: the modeling approach and parameter sensitivity analysis.
    \item \textbf{Revised manuscript}: The updated and final version of the manuscript that incorporates changes based on the reviewers' feedback and our own revisions.
    \item \textbf{Tracked changes summary}: A document showing all modifications made to the initial submission, with color-coded text and explanations for added, removed, and revised content.
\end{itemize}

% \begin{enumerate}
%     \item \textbf{Modeling approach}: 
    
%     We recognize that the initial submission provided a limited explanation of the original model. To address this, we have expanded our explanation, while keeping the changes to the model minimal. These revisions directly address the reviewers' concerns and remain consistent with the intent of the initial model.

%     Our work aims to establish a foundation for future research in this area by developing a modeling and control strategy within a general framework. This framework is designed to encompass key features of real-world systems, rather than focusing on a model tailored to a specific system. 

%     \item \textbf{Parameter sensitivity analysis}: We noted the reviewers' interest in parameter analysis. While we tested various parameter sets during our initial simulations, we did not include them in the original manuscript to maintain focus. The varied comments from reviewers highlight how broad this topic can be. However, as state delay is central to our work, we have added an analysis showing how the controller’s performance changes with different delay parameters.
% \end{enumerate}

Thank you again for considering our work. We do hope the revised manuscript meets the expectations of the reviewers and the journal.

\closing{Sincerely,}

\vspace{-4em} % Adjust the space between closing and signature

Behrad Moadeli \\
Department of Chemical and Materials Engineering \\
University of Alberta, Edmonton, Alberta, Canada \\
moadeil@ualberta.ca

\end{letter}

\end{document}
