\documentclass[letterpaper,11pt]{article}

% Preamble
\usepackage[utf8]{inputenc}
\usepackage{amsmath, amssymb}
\usepackage{graphicx}
\usepackage{float}
\usepackage{hyperref}
\usepackage{subcaption}

\usepackage[style=chem-acs]{biblatex}
\addbibresource{references.bib}
\usepackage{geometry}
\geometry{letterpaper, margin=1in}

\usepackage{tikz}
\usetikzlibrary{arrows.meta, positioning}  % Load arrows.meta for 'latex' arrow tip

\usepackage{setspace}
\doublespacing

\usepackage{titlesec}
\titleformat{\subsubsection}{\normalfont\normalsize}{\thesubsubsection}{1em}{}

\usepackage{caption}
\renewcommand{\figurename}{\textbf{\uppercase{Figure}}}
\renewcommand{\thefigure}{\arabic{figure}}
\captionsetup[figure]{
    labelsep=colon,
    labelfont={bf,small}, % Make label bold in caption
    textfont={small}, % Text in caption remains unbolded and small
    labelformat=default
}
\makeatletter
\let\oldthefigure\thefigure
\renewcommand{\thefigure}{\arabic{figure}}
\makeatother


\captionsetup[subfigure]{
    labelformat=parens,       % Add parentheses around the label
    labelsep=space,           % Add a space between the label and the caption
    textfont=normalfont,      % Use normal font for the caption text
    position=bottom,          % Position the label at the bottom of the subfigure
    % labelformat=simple,       % Use a simple format without automatic parentheses
    format=plain              % Plain format for subfigure captions
}
\renewcommand\thesubfigure{\Alph{subfigure}}


\usepackage{lipsum}
\usepackage{multicol}


% Title and author information
\title{Optimal Control of Axial Tubular Reactors with Recycle Streams:\\Addressing State-Delay through Transport PDEs}
\author{
  Behrad Moadeli \thanks{Department of Chemical \& Materials Engineering, University of Alberta, Edmonton, Alberta, Canada. \\Email: moadeli@ualberta.ca} \and
  Guilherme Ozorio Cassol \thanks{Department of Chemical \& Materials Engineering, University of Alberta, Edmonton, Alberta, Canada. \\Email: ozorioca@ualberta.ca} \and
  Stevan Dubljevic \thanks{Department of Chemical \& Materials Engineering, University of Alberta, Edmonton, Alberta, Canada. \\Email: stevan.dubljevic@ualberta.ca}
}

\date{\today}

\begin{document}

\maketitle

\begin{abstract}
    % Your abstract here
    The optimal control of an axial tubular reactor with a recycle stream is addressed as a key class of distributed parameter systems in chemical engineering. The intrinsic time delay from the recycle process, thus far overlooked in relevant literature, is modeled as a transport partial differential equation (PDE), resulting in a system of coupled parabolic and hyperbolic PDEs. Utilizing Danckwerts boundary conditions, the reactor is boundary-controlled with the control input at the inlet. A continuous-time optimal linear quadratic regulator is developed to stabilize this infinite-dimensional system, employing a late lumping approach in order to preserve the properties of the original infinite dimensional system in controller design. The full-state feedback regulator is designed by solving the Operator Riccati Equation (ORE), leveraging the system's Riesz-spectral properties. To address practical limitations of full-state feedback, a Luenberger observer is also proposed, enabling state reconstruction from boundary measurements. 
    Numerical simulations are conducted to evaluate the proposed control strategies. The results demonstrate that the full-state feedback regulator effectively stabilizes the system. A comparison is made between two configurations where different numbers of eigenmodes were selected to design the controller. The observer-based regulator, though slightly slower, also successfully stabilizes the system using only output measurements, effectively overcoming the challenge of limited state access.
\end{abstract}

\newpage
\tableofcontents
\newpage

% \begin{multicols}{2}

  \input{Sections/01_intro.tex}
  
  \input{Sections/02_model.tex}
  
  \input{Sections/03_control.tex}
    
  \input{Sections/04_results.tex}

  \input{Sections/05_conc.tex}
  
  \newpage
  \printbibliography

  % \end{multicols}

\input{Figures/Figures.tex}

\end{document}
